Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics

### 2017-08-29 Manuscripts_GMD_ap.tex
new master tex

%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
% \Author[affil]{given_name}{surname}
\Author[1]{Zhixuan}{Cao}
\Author[1]{Abani}{Patra}
\Author[2]{Marcus}{Bursik}
\Author[3]{E. Bruce}{Pitman}
\Author[4]{Matthew}{Jones}
\affil[1]{Department of Mechanical and Aerospace Engineering, University at Buffalo, SUNY, New York, United State}
\affil[2]{Department of Geology, University at Buffalo, SUNY, New York, United State}
\affil[3]{Department of Material Design and Innovation, University at Buffalo, SUNY, New York, United State}
\affil[4]{Center for Computational Research,
University at Buffalo, SUNY, New York, United State}
%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
\runningtitle{Plume-SPH 1.0}
\runningauthor{Z. Cao et al}
%<<<<<<< HEAD
\correspondence{Abani Patra (abani@buffalo.edu)}
%=======
%\correspondence{Abani Patra (abani.patra@gmail.com)}
%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}
%% These dates will be inserted by Copernicus Publications during the typesetting process.
\firstpage{1}
\maketitle
\begin{abstract}
VATDs (Volcanic ash transport and dispersion models) used for volcanic ash forecasts often take outputs of eruption plume models as source terms. The accuracy of these terms is crucial for forecasts from VATDs. All existing 3D (three dimensional) plume models described in the literature use mesh based methods. SPH (smoothed particle hydrodynamics), as a mesh free method, has several advantages over mesh based methods in modeling of multiphase free boundary flows. As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple model, a 3D dusty-gas dynamic model, targeted at capturing the salient features of a volcanic plume. In this model, erupted material is assumed to be well-mixed, and represented by one phase while air is another phase. We also assume dynamic and thermodynamic equilibrium between air and erupted material and that the effect of wind is not significant.

Several newly developed techniques in SPH are adapted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method. The $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales. Heat exchange due to turbulence is then calculated by Reynolds analogy. A corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries. We also propose to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH. The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability. The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment). Next, the top height of the Pinatubo eruption of 15 June 1991, as simulated by our model, is consistent with both observations and existing 3D plume models. Profiles of several integrated variables are compared with those calculated in existing 3D plume models, and further verify our model. Analysis of the plume evolution process illustrates that this model is able to reproduce the basic physics of plume development. The comparison also implies that the turbulence model plays a significant role in 3D volcanic plume modeling.
\end{abstract}

@@

One of the most commonly used models of artificial viscosity is:
\begin{equation}
\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2}
\label{eq:art-vis-original}
\end{equation}
The coefficient $\nu$ is defined as:
\begin{equation}
\nu = \alpha \bar{h}_{ab} \bar{c}_{ab} \label{eq:def_nu_in_av}
\end{equation}
Where 
\begin{align}
\bar{c}_{ab} = \dfrac{c_a + c_b}{2} \\
\bar{\rho}_{ab} = \dfrac{\rho_a + \rho_b}{2} \\
\textbf{v}_{ab}=\textbf{v}_a-\textbf{v}_b \\
\textbf{x}_{ab}=\textbf{x}_a-\textbf{x}_b
\end{align}
The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation. It produces a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.

@@

\subsection{Turbulence modeling with SPH}
Typically, turbulence fluctuations in a volcanic column occurs at very different length scales, ranging from millimeters up to hundreds of meters. The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical exchange coefficients (molecular viscosity and heat conduction coefficient). One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model. 
Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant–Friedrichs–Lewy) condition. Here we choose the SPH-SPS method. Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}. However, the $SPH-\varepsilon$ turbulence model was proposed only for incompressible flow. In the following section, we will extend it for compressible flow. It is necessary to mention that all other existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical} also only focus on incompressible flow.

\subsubsection{Langrangian average in $SPH-\varepsilon$}
\citet{monaghan2011turbulence} constructed the $SPH-\varepsilon$ turbulence model within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS turbulence modeling. The basic idea of $SPH-\varepsilon$ is to determine a smoothed (averaged) velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$. The SPH particles move with this smoothed velocity and hence the average motion of the fluid is determined by the averaged velocity $\widehat{\textbf{v}}$:
\begin{equation}
\dfrac{d \textbf{x}_a}{dt} = \widehat{\textbf{v}}_a \label{eq:gov-update-pos-turbulence}
\end{equation}

@@

It is a common practice in LANS to use a differential equation for the smoothing rather than the integral form and finally reach to a system of equations that need to be solved implicitly. In $SPH-\varepsilon$ method, a XSPH \citep{monaghan1989problem} smoothing is adopted which conserves linear and angular momentum. In this way, solving of a system of equations is avoided and hence makes the method simple to implement and cheap for computation. 
The discretized form of the momentum equation is obtained through lengthy derivation. Derivation and other discussions are available in the literature -- see for e.g. \citep{monaghan2011turbulence}. Here we provide a brief summary of key steps.

The smoothing adopted by \citet{monaghan2011turbulence}  is:
\begin{equation}
\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+ \epsilon \int \left(\textbf{v}\left(\textbf{x} \prime\right)-\textbf{v}\left(\textbf{x}\right)\right)G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime
\end{equation}
As function $G$ has the same feature as kernel function $w$, SPH approximation of the integration leads to:
\begin{equation} \label{eq:SPH-epsilon-filtering}
\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+\epsilon \sum_b m_b \dfrac{\left(\textbf{v}_b -\textbf{v}\right)}{\rho _b} G\left(\vert \textbf{x} _b - \textbf{x} \vert, l\right)
\end{equation}
By making the replacement:
\begin{equation}
\label{eq:replacement-in-turb-derive}
\dfrac{G\left(\vert \textbf{x} _b - \textbf{x} _a \vert, l\right)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
\end{equation}
Where $K_{ab} = l^d G_{ab}$, $M = \rho_b l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
\begin{equation}
\label{eq:monaghan-mom-turb}
\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
\end{equation}
Notice that if $l$ is uniform: 
%Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq. (2.17)). The original equation is based on the assumption that density is uniform initially. This assumption is not general and atmosphere density is stratified in our plume simulation. And the slight modification will not influence following derivation if $l$ is uniform and keep constant because:
%
%{\bf WE NEED TO SAY SOMETHING ABOUT INCOMPRESSIBILITY HERE}
%
\begin{equation}
\nabla K_{ab} = \nabla \left(l^d G_{ab}\right) = l^d \nabla G_{ab}
\end{equation}
The discretized momentum equation with $SPH-\varepsilon$ turbulence model can be written in terms of $G_{ab}$ instead of $K_{ab}$:
\begin{equation}
\label{eq:SPH-mom-epsilon-turb}
\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}\left(l\right)
\end{equation}
where 
\begin{equation}
\Phi_{ab}=\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b} \label{eq:_LANS_trub_stress}
\end{equation}
Which is the extra stress term induced by average. We take coefficient $\varepsilon$ as 0.8 following \citet{monaghan2011turbulence}.

For compressible flow, the energy conservation equation is coupled with the momentum conservation equation and mass conservation equation. Averaging of thermal energy over space introduces some additional terms besides the stress term induced by velocity average. \citep{NASACompressibleTurbulence}. The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms, besides the corresponding velocity average induced stress term, show up in the energy equation. Most turbulence modeling focuses on the stress terms induced by average of velocity. These stress terms are usually either solved directly (for example, LANS methods) or defined via a constitutive relation (for example, large eddy simulation method). Less attention is typically given to other terms. Most commonly, a Reynolds analogy is used to model the turbulent exchange. Simulations of heat transfer, or other scalar transfer due to turbulence simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}. We adopt this strategy. %PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}. We neglect these terms in our simulation.

\subsubsection{Turbulent heat transfer}
We adopt Reynolds analogy to get the heat transfer coefficient due to turbulence.
The Prandtl number is defined as:
\begin{equation}
Pr=\dfrac{C_p \mu}{\kappa}
\end{equation}
Where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity).
\begin{align}
\mu=\rho \nu
\end{align}
Then
\begin{equation}
\kappa=\dfrac{C_p \mu}{Pr}
\end{equation}
Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended \citet{kays1994turbulent} from summarizing of experimental results.

@@

However, the above equation is correct only for the 1D situation. For 2D or 3D, it is not easy to get an explicit expression. 
We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression. Choosing the smoothing function as the same as SPH kernel and the smoothing length scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is: 
\begin{equation}
\begin{split}
\Upsilon_{ab} &= \dfrac{\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b}}{\dfrac{S \nu}{\rho_{ab}} \frac{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2 + \eta^2 h_{ab}^2}} \\
 & = \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S \nu} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{split}
\end{equation}
$\Upsilon_{ab}$ is essentially equivalent to the ratio between turbulent viscous effect of particle b on particle a and molecular viscous effect of particle b on particle a. Turbulent viscosity can be easily obtained by:
\begin{equation}
\begin{split}
\nu_{t,ab} &= \nu \Upsilon_{ab} \\
&= \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{split}
\end{equation}
Recall that in Eq. (\ref{eq:def_nu_in_av}), $\nu$ is scaled by $\bar{h}_{ab} \bar{c}_{ab}$. Then the corresponding turbulent thermal conductivity should be:
\begin{equation}
\kappa_{t,ab}=\dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} \left(x_{ab}^2 + \eta^2 h_{ab}^2\right) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{equation}
$\bar{C}_{p,ab}$ and $\bar{\rho}_{ab}$ are simply the arithmetic means of specific heat and density. The term used to prevent singularity can be removed. 
\begin{equation}
\kappa_{t,ab}=\dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab} }
\end{equation}
We also need to prevent singularity and guarantee positive conductivity coefficient, so: 
\begin{equation}
\kappa_{t,ab}= 
\begin{cases} 
      0 & if \quad \textbf{v}_{ab} \cdot \textbf{x}_{ab} \leq 0 \\
      \dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab} } & \text{otherwise}
\end{cases}
\label{eq:SPH-LANS-heat-conductivity}
\end{equation}

>>>>

%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
% \Author[affil]{given_name}{surname}
\Author[1]{Zhixuan}{Cao}
\Author[1]{Abani}{Patra}
\Author[2]{Marcus}{Bursik}
\Author[3]{E. Bruce}{Pitman}
\Author[4]{Matthew}{Jones}
\affil[1]{Department of Mechanical and Aerospace Engineering and Computational Data Science and Engineering, University at Buffalo, SUNY, New York, USA}
\affil[2]{Department of Geology, University at Buffalo, SUNY, New York, USA}
\affil[3]{Department of Material Design and Innovation, University at Buffalo, SUNY, New York, USA}
\affil[4]{Center for Computational Research,
University at Buffalo, SUNY, New York, USA}
%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
\runningtitle{Plume-SPH 1.0}
\runningauthor{Z. Cao et al}
%<<<<<<< HEAD
\correspondence{Abani Patra (abani@buffalo.edu)}
%=======
%\correspondence{Abani Patra (abani.patra@gmail.com)}
%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}
%% These dates will be inserted by Copernicus Publications during the typesetting process.
\firstpage{1}
\maketitle
\begin{abstract}

Plume-SPH provides the the first particle based simulation of volcanic plumes.  SPH (smoothed particle hydrodynamics) has several advantages over currently used mesh based methods in modeling of multiphase free boundary flows like volcanic plumes. This tool will provide more accurate eruption source terms to users of VATDs (Volcanic ash transport and dispersion models) greatly improving volcanic ash forecasts.  The accuracy of these terms is crucial for forecasts from VATDs and the 3D SPH model presented here will provide better numerical accuracy. As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model  (3D dusty-gas dynamic model assuming well mixed eruption materiel and dynamic and thermodynamic equilibrium between air and erupted material and minimal effect of winds) targeted at capturing the salient features of a volcanic plume. The documented open source code  is easily obtained and extended to incorporate other models of physics of interest to the large community of researchers investigating multiphase free boundary flows of volcanic or other origins.

The Plume-SPH code also incorporates  several newly developed techniques in SPH  needed to address numerical challenges in simulating multiphase compressible turbulent flow.  The code should thus be also of general interest to the much larger community of researchers using and developing SPH based tools. In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales, heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries. We also developed methodology to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.

The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability using novel techniques for data management using a SFCs (space-filling curves) and object creation time based indexing and hash table based storage scheme. These techniques are of  interest to  researchers engaged in developing particle in cell type methods. The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991. Our results are consistent with both observations and existing 3D plume models. Profiles of several integrated variables are compared with those calculated in existing 3D plume models, and further verify our model. Analysis of the plume evolution process illustrates that this model is able to reproduce the physics of plume development.

\end{abstract}

@@

One of the most commonly used models of artificial viscosity is:
\begin{equation}
\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2}
\label{eq:art-vis-original}
\end{equation}
The coefficient $\nu$ is defined as:
\begin{equation}
\nu = \alpha \bar{h}_{ab} \bar{c}_{ab}
\end{equation}
Where 
\begin{align}
\bar{c}_{ab} = \dfrac{c_a + c_b}{2} \\
\bar{\rho}_{ab} = \dfrac{\rho_a + \rho_b}{2} \\
\textbf{v}_{ab}=\textbf{v}_a-\textbf{v}_b \\
\textbf{x}_{ab}=\textbf{x}_a-\textbf{x}_b
\end{align}
The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation. It produces a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.

@@

\subsection{Turbulence modeling with SPH}
Typically, turbulence fluctuations in a volcanic column occurs at very different length scales, ranging from millimeters up to hundreds of meters. The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical coefficients (molecular viscosity and heat conduction coefficient). One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model. 
Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant–Friedrichs–Lewy) condition. Here we choose the SPH-SPS method. Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}. However, the $SPH-\varepsilon$ turbulence model was proposed only for incompressible flow. In the following section, we will extend it for compressible flow. It is necessary to mention that all other existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical} also only focus on incompressible flow.

%Even though free surface flow are in nature turbulent, detailed research on turbulence modeling in SPH method is rather scarce. Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation, monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach \citep{violeau2007numerical}. The first wall-bounded LES (large eddy simulation) simulation with SPH was performed by \citet{issa2005numerical}. An improved version of $SPH-\alpha$ method  \citep{monaghan2002sph} is proposed by \citet{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method. We adopt $SPH-\varepsilon$ method in our model. Most of these papers constrain their focus to incompressible flow. 
%$SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion. However, in that paper, energy conservation equation is discretized in a way that is exactly the same as usual SPH. There is no special treatment on energy conservation equation. The recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focuses on incompressible flow only.
%

\subsubsection{Langrangian average in $SPH-\varepsilon$}
\citet{monaghan2011turbulence} constructed $SPH-\varepsilon$ turbulence model within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS turbulence modeling. The basic idea of $SPH-\varepsilon$ is to determine a smoothed (averaged) velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$. The SPH particles move with this smoothed velocity and hence the average motion of the fluid is determined by the averaged velocity $\widehat{\textbf{v}}$:
\begin{equation}
\dfrac{d \textbf{x}_a}{dt} = \widehat{\textbf{v}}_a \label{eq:gov-update-pos-turbulence}
\end{equation}

@@

It is a common practice in LANS to use a differential equation for the smoothing rather than the integral form and finally reach to a system of equations that need to be solved implicitly. In $SPH-\varepsilon$ method, a XSPH \citep{monaghan1989problem} smoothing is adopted which conserves linear and angular momentum. In this way, solving of a system of equations is avoided and it also makes the method simple to implement and cheap for computation. 
The discretized form of the momentum equation is obtained through lengthy derivation. Derivation and other discussions are available in the literature -- see for e.g. \citep{monaghan2011turbulence}. Here we provide a brief summary of key steps.

The smoothing adopted by \citet{monaghan2011turbulence}  is:
\begin{equation}
\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+ \epsilon \int \left(\textbf{v}\left(\textbf{x} \prime\right)-\textbf{v}\left(\textbf{x}\right)\right)G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime
\end{equation}
As function $G$ has the same feature as kernel function $w$, SPH approximation of the integration leads to:
\begin{equation} \label{eq:SPH-epsilon-filtering}
\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+\epsilon \sum_b m_b \dfrac{\left(\textbf{v}_b -\textbf{v}\right)}{\rho _b} G\left(\vert \textbf{x} _b - \textbf{x} \vert, l\right)
\end{equation}
By making the replacement:
\begin{equation}
\label{eq:replacement-in-turb-derive}
\dfrac{G\left(\vert \textbf{x} _b - \textbf{x} _a \vert, l\right)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
\end{equation}
Where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
\begin{equation}
\label{eq:monaghan-mom-turb}
\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
\end{equation}
Notice that if $l$ is uniform: 
%Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq. (2.17)). The original equation is based on the assumption that density is uniform initially. This assumption is not general and atmosphere density is stratified in our plume simulation. And the slight modification will not influence following derivation if $l$ is uniform and keep constant because:
%
%{\bf WE NEED TO SAY SOMETHING ABOUT INCOMPRESSIBILITY HERE}
%
\begin{equation}
\nabla K_{ab} = \nabla \left(l^d G_{ab}\right) = l^d \nabla G_{ab}
\end{equation}
The discretized momentum equation with $SPH-\varepsilon$ turbulence model can be written in terms of $G_{ab}$ instead of $K_{ab}$:
\begin{equation}
\label{eq:SPH-mom-epsilon-turb}
\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}\left(l\right)
\end{equation}
where 
\begin{equation}
\Phi_{ab}=\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b} 
\end{equation}
Which is the extra stress term induced by average. We take coefficient $\varepsilon$ as 0.8 following \citet{monaghan2011turbulence}.

For compressible flow, the energy equation is coupled with the momentum equation and mass conservation equation. Averaging of thermal energy over space introduces some additional terms besides the stress term induced by velocity average. \citep{NASACompressibleTurbulence}. The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms, besides the corresponding velocity average induced stress term, show up in the energy equation. Most turbulence modeling focuses on the stress terms induced by average of velocity. These stress terms are usually either solved directly (for example, LANS methods) or defined via a constitutive relation (for example, large eddy simulation method). Less attention is typically given to the other terms. Most commonly, a Reynolds analogy is used to model the turbulent exchange. Simulations of heat transfer, or other scalar transfer, in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}. We adopt this strategy. %PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}. We neglect these terms in our simulation.

\subsubsection{Turbulent heat transfer}
We adopt Reynolds analogy to get the heat transfer coefficient due to turbulence.
The Prandtl number is defined as:
\begin{equation}
Pr=\dfrac{C_p \mu}{\kappa}
\end{equation}
Where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity) as:
\begin{align}
\mu=\rho \nu
\end{align}
Then
\begin{equation}
\kappa=\dfrac{C_p \mu}{Pr}
\end{equation}
Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended \citet{kays1994turbulent} from summarizing of experimental results.

@@

However, the above equation is correct only for the 1D situation. For 2D or 3D, it is not easy to get an explicit expression. 
We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression. Choosing the smoothing function as the same as SPH kernel and the smoothing length scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is: 
\begin{equation}
\begin{split}
\Upsilon_{ab} &= \dfrac{\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b}}{\dfrac{S \nu}{\rho_{ab}} \frac{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2 + \eta^2 h_{ab}^2}} \\
 & = \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S \nu} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{split}
\end{equation}
$\Upsilon_{ab}$ is essentially equivalent to the ratio between turbulent viscous effect of particle b on particle a and molecular viscous effect of particle b on particle a. Turbulent viscosity can be easily obtained by:
\begin{equation}
\begin{split}
\nu_{t,ab} &= \nu \Upsilon_{ab} \\
&= \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{split}
\end{equation}
The corresponding turbulent thermal conductivity should be
\begin{equation}
\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} \left(x_{ab}^2 + \eta^2 h_{ab}^2\right) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{equation}
$\overline{C_{p,ab}}$ and $\overline{\rho_{ab}}$ are simply the arithmetic means of specific heat and density. The term used to prevent singularity now can be removed. 
\begin{equation}
\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab} }
\end{equation}
We also need to prevent singularity, so: 
\begin{equation}
\kappa_{t,ab}= 
\begin{cases} 
      0 & if  \quad \textbf{v}_{ab}=0 \quad or \quad \textbf{x}_{ab}=0 \\
      \dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab} } & \text{otherwise}
\end{cases}
\label{eq:SPH-LANS-heat-conductivity}
\end{equation}

Heat conduction equation without source term is:
\begin{equation}
C_p \dfrac{dT}{dt} = \dfrac{1}{\rho} \nabla \left(\kappa \nabla T\right)
\end{equation}
Second spatial derivative can be approximated with SPH by following \citet{monaghan2005smoothed}. 
\begin{equation}
C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} \left(\kappa_a + \kappa_b\right) \left(T_a - T_b\right) F_{ab} \left(h\right)
\end{equation}
Where
\begin{equation}
F_{ab} \textbf{x}_{ab} = \nabla _a w_{ab}
\end{equation}
$F_{ab}$ is nonpositive, which guarantees that heat flux flows from hot to cold. 
Plug the turbulent thermal conductivity into the heat conduction equation:
\begin{equation}
\begin{split}
C_p \dfrac{dT}{dt}
& = \sum_b \dfrac{m_b}{\rho_a \rho_b} \left(\kappa_a + \kappa_b\right) \left(T_a - T_b\right) F_{ab} \left(h\right) \\
 &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } \left(T_a - T_b\right) F_{ab} \left(h\right)
\end{split}
\label{eq:turb-LANS-heat-conduct1}
\end{equation}
Notice that the number "2" in the front of Eq. (\ref{eq:turb-LANS-heat-conduct1}) comes from integration approximation of second order derivative \citep {cleary1999conduction}. By further simplification, we get:
\begin{equation}
C_p \dfrac{dT}{dt}
 =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} \left(T_a - T_b\right) F_{ab} \left(h\right)
\end{equation}
